PineAPPL: EW and mixed QCD--EW corrections for PDF fits

### 2020-07-30 paper/sec-results.tex
Sect 3

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme for all massive particles, see
Ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that do not involve stable top quarks in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: PI effects and EW corrections
are therefore treated on the same footing. We use a PDF set that contains a
consistently defined photon PDF, namely
{\tt NNPDF31\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation.

The central values of the renormalisation and factorisation scales, $\mu_R$ and
$\mu_F$, are chosen, for each process, as follows. In the case of DY lepton pair
production, we use the fixed scale $\mu_R=\mu_F=M_Z$, where $M_Z$ is the mass
of the $Z$-boson, for the ATLAS measurement, and the scale
$\mu_R=\mu_F=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS measurement.
In the case of top-quark pair production, we use the dynamic scales
$\mu_R=\mu_F=\sqrt{m_t^2+(p_T^t)^2}{\Big /}2$ for the distribution differential
in the transverse momentum of the top quark, and $\mu_R=\mu_F=H_T/4$ for the
distribution differential in the invariant mass of the top-quark pair, where
$H_T=\sqrt{m_t^2+(p_T^t)^2}+\sqrt{m_t^2+(p_T^{\bar{t}})}$, with $m_t$,
$p_T^t$ and $p_T^{\bar t}$ the mass of the top quark and the transverse momenta
of the top and antitop quarks, respectively. These choices were demonstrated
to maximise the convergence of the perturbative expansion~\cite{Czakon:2016dgf}.
In the case of $Z$-boson production with non-zero transverse momentum, we use
$\mu_R=\mu_F=M_Z$. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted in the Monte Carlo generation upon scale
variations. As is customary, the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and an envelope
from the nine-point scale variations is constructed.

@@

For each of the measurements discussed in
Sect.~\ref{subsec:processes_and_measurements}, we compute the expectation
value of the corresponding observable for each kinematic bin in two different
ways: directly, by means of {\sc MG5\_aMC}, and a posteriori, by convolving the
fast interpolation grid produced by {\sc PineAPPL} with the PDF set specified
in Sect.~\ref{subsec:computational_settings}. In the following, we will refer
to the the first result as the {\sc MC} result, and to the second as the
{\sc PineAPPL} result. We repeat the computation for theories accurate to NLO
QCD and to NLO QCD+EW, respectively. In each case, we determine the PDF
uncertainty (coming from the PDF ensemble), the scale uncertainty (coming from
variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of a fraction of permille. This choice
ensures that the statistical uncertainty of the computation is negligible
in comparison to the PDF and scale uncertainties, as we will explicitly
demonstrate.
%Add here a sentence on a low-statistics run?

Our goal is twofold. On the one hand, we aim to validate the fast interpolation
grids generated with {\sc PineAPPL}: to this purpose we shall verify that the
MC and the {\sc PineAPPL} results are identical up to numerical inaccuracies
due to the grid interpolation. This equivalence must hold for any choice of
renormalisation and factorisation scale. On the other hand, we aim to study
the size of the EW corrections, in particular with respect to the kinematics of
each process, and to an array of relative uncertainties arising from the PDFs,
the scale variations, and the experimental data.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLAS_TTB_DIFF_8TEV_LJ_TPT}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLAS_TTB_DIFF_8TEV_LJ_TTM}
    \caption{PineAPPL comparison for ATLAS top pair.}
    \label{fig:cmsdy2d11_bins56}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMS_Z_13_TEV}
    \caption{PineAPPL comparison for CMS $Z$ $p_T$ distribution.}
    \label{fig:cmsdy2d11_bins56}
\end{figure}

\noindent
TODO for each of the following subsections:
\begin{itemize}
\item size of the photon-initiated contributions,
\item largest partonic channel,
\item most important $x$ region,
\item PDF uncertainty,
\end{itemize}

%ERN 7 Apr: there is consensus on the fact that we should present results for the
%following data sets:
%\begin{itemize}
%\item ATLAS high mass DY distributions, 7 TeV~\cite{Aad:2013iua} (CS);
%\item CMS 2D DY distributions, 7 TeV~\cite{Chatrchyan:2013tia} (CS);
%\item ATLAS top pair differential distributions ($m_{t\bar{t}}$ and $p_T^t$),
%8 TeV~\cite{Aad:2015mbv} (ERN);
%\item CMS $Z$ $pT$ distributions, 13 TeV~\cite{Sirunyan:2019bzr} (ERN).
%\end{itemize}
%
%We agree not to display any LHCb measurement, given that they won't add
%further value to our discussion.
%Note added: we might also want to have a look at the ATLAS 2D and 3D DY
%distributions, 8 TeV~\cite{Aad:2016zzw,Aaboud:2017ffb}, if time allows
%us to do so.
%
%CS 25 Jun: We've agreed to show basically two types of plots: 1) technical plots showing the good agreement between the MC compared to the results from the grids, and 2) phenomenological results showing larger EW corrections, for example.
%In the aMCfast paper both is shown in single plot, but since we have more to show, I suggest the following: for the technical plots we show a 2x2 matrix of plots, each showing the difference of the grid result compared to the MC result, in the following fashion:
%\begin{itemize}
%\item NLO QCD with low statistics,
%\item NLO QCD+EW with low statistics,
%\item NLO QCD with high statistics, and finally
%\item NLO QCD+EW with high statistics,
%\end{itemize}
%each showing a few scale variations.
%These plots then clearly show that no matter what corrections you choose, no matter the statistics, and no matter the scale variation, the agreement is always excellent.
%In any case I would like to avoid showing a plot with absolute numbers and low statistics, which looks a bit ridiculous in my opinion (look at figure 1, left side, top plot of the aMCfast paper).
%
%Finally we can show another series of plots, which in my opinion should be very similar to the usual pheno paper plots: absolute numbers with a scale variation band, maybe a few corrections shown in the same plot and then in the bottom relative corrections.

>>>>

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme for all massive particles, see
Ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that do not involve stable top quarks in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: PI effects and EW corrections
are therefore treated on the same footing. We use a PDF set that contains a
consistently defined photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation.

The central values of the renormalisation and factorisation scales, $\mu_R$ and
$\mu_F$, are chosen, for each process, as follows. In the case of DY lepton pair
production, we use the fixed scale $\mu_R=\mu_F=M_Z$, where $M_Z$ is the mass
of the $Z$-boson, for the ATLAS measurement, and the scale
$\mu_R=\mu_F=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS measurement.
In the case of top-quark pair production, we use the dynamic scales
$\mu_R=\mu_F=\sqrt{m_t^2+(p_T^t)^2}{\Big /}2$ for the distribution differential
in the transverse momentum of the top quark, and $\mu_R=\mu_F=H_T/4$ for the
distribution differential in the invariant mass of the top-quark pair, where
$H_T=\sqrt{m_t^2+(p_T^t)^2}+\sqrt{m_t^2+(p_T^{\bar{t}})}$, with $m_t$,
$p_T^t$ and $p_T^{\bar t}$ the mass of the top quark and the transverse momenta
of the top and antitop quarks, respectively. These choices were demonstrated
to maximise the convergence of the perturbative expansion~\cite{Czakon:2016dgf}.
In the case of $Z$-boson production with non-zero transverse momentum, we use
$\mu_R=\mu_F=M_Z$. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted in the Monte Carlo generation upon scale
variations. As is customary, the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and an envelope
from the seven-point scale variations is constructed.

@@

For each of the measurements discussed in
Sect.~\ref{subsec:processes_and_measurements}, we compute the expectation
value of the corresponding observable for each kinematic bin in two different
ways: directly, by means of {\sc MG5\_aMC}, and a posteriori, by convolving the
fast interpolation grid produced by {\sc PineAPPL} with the PDF set specified
in Sect.~\ref{subsec:computational_settings}. In the following, we will refer
to the the first result as the {\sc MC} result, and to the second as the
{\sc PineAPPL} result. We repeat the computation for theories accurate to NLO
QCD and to NLO QCD+EW, respectively. In each case, we determine the PDF
uncertainty (coming from the PDF ensemble), the scale uncertainty (coming from
variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of a fraction of permille. While
this choice does not affect the validation of the {\sc PineAPPL} result against
the {\sc MC} result, it ensures that the statistical uncertainty of the
computation remains negligible in comparison to the PDF and scale uncertainties,
as we will explicitly demonstrate. This is a desirable feature to correctly
interpret the size of the theoretical corrections.

Our goal is indeed twofold. On the one hand, we aim to validate the
interpolation grids generated with {\sc PineAPPL}: to this purpose we shall
verify that the MC and the {\sc PineAPPL} results are identical up to numerical
inaccuracies due to the grid interpolation. This equivalence must hold for any
choice of renormalisation and factorisation scale. On the other hand, we aim to
study the size of the EW corrections, in particular with respect to the
kinematics of each process, and to three kinds of uncertainties: the PDF
uncertainty, the scale variation uncertainty, and the uncertainty of the
experimental data.

We present these comparisons in
Figs.~\ref{fig:atlaszhighmass49fb}-\ref{fig:cmsZ13TeV} for each of the processes
and data sets outlined in Sect.~\ref{subsec:processes_and_measurements}.
The format of the plots is the same across all figures. The first panel
displays the relative difference (in permille) between the {\sc PineAPPL} and
the {\sc MC} results for the central, upper and lower scale choices, for
theories accurate to both NLO QCD and NLO QCD+EW. The following three panels
present the theoretical predictions, accurate to either NLO QCD or NLO QCD+EW,
always normalised to the former: the second displays the PDF uncertainty, the
third the scale uncertainty and the fourth the Monte Carlo statistical
uncertainty. The relative uncertainty of the experimental data points is also
shown for comparison. We shall now discuss the results for each
process and data set in turn.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLAS_TTB_DIFF_8TEV_LJ_TPT}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLAS_TTB_DIFF_8TEV_LJ_TTM}
    \caption{PineAPPL comparison for ATLAS top pair.}
    \label{fig:atlastop}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMS_Z_13_TEV}
    \caption{PineAPPL comparison for CMS $Z$ $p_T$ distribution.}
    \label{fig:cmsZ13TeV}
\end{figure}

\noindent

TODO for each process:
\begin{itemize}
\item size of the photon-initiated contributions,
\item largest partonic channel,
\item most important $x$ region,
\item PDF uncertainty,
\end{itemize}
